---
title: "Untitled"
author: "Marina Sanchez"
date: "11/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#install.packages("dplyr")
#install.packages("viridis")  
library("viridis")
library(dplyr)
library(tidyverse)
library(ggpubr)
library(Matrix)
library(qdapDictionaries)
library(readxl)
library(ggplot2)
```


```{r}

#read in data and check for NaNs
df <- read.csv("spotify_tracks.csv")

#apply(df, 2, function(x) sum(is.na(x)))
apply(df, 2, function(x) sum(is.finite(x)))
apply(df, 2, function(x) sum(is.na(x) | x == ''))

```

```{r}
## Scatterplots for infinite variables vs popularity

tracks.df <- df[,c("acousticness", "danceability", "available_markets", "country", "duration_ms", "energy", "instrumentalness", "liveness", "loudness", "lyrics", "speechiness", "valence", "key", "tempo", "popularity", "id")]


# data transformations
tracks.df$popularity <- tracks.df$popularity/100
tracks.df$duration_scaled <- tracks.df$duration_ms/5505831
tracks.df$loudness_scaled <- (tracks.df$loudness + 60)/62.719

# display r correlations on plots
upper.panel <- function(x,y){
  points(x,y, pch=1)
  r <- round(cor(x,y), digits = 2)
  txt <- paste0("R = ", r)
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  text(0.5, 0.9, txt)
}

# scatter plots - arranged 3 x 3 for clarity
pairs(tracks.df[c(15, 1:2)], lower.panel = NULL,
      upper.panel = upper.panel)

tracks.df <- tracks.df[,c("acousticness", "danceability", "available_markets", "country", "duration_scaled", "energy", "instrumentalness", "liveness", "loudness_scaled", "lyrics", "speechiness", "valence", "key", "tempo", "popularity", "id")]


##Plots below not to be included - for reference only
# display r correlations on plots
# upper.panel <- function(x,y){
#   points(x,y, pch=1)
#   r <- round(cor(x,y), digits = 2)
#   txt <- paste0("R = ", r)
#   usr <- par("usr"); on.exit(par(usr))
#   par(usr = c(0, 1, 0, 1))
#   text(0.5, 0.9, txt)
# }

# scatter plots - arranged 3 x 3 for clarity
# pairs(tracks.df[c(15, 1:2)], lower.panel = NULL,
#       upper.panel = upper.panel)
# 
# pairs(tracks.df[c(15, 13:14)], lower.panel = NULL,
#       upper.panel = upper.panel)
# 
# pairs(tracks.df[c(15, 5:6)], lower.panel = NULL,
#       upper.panel = upper.panel)
# 
# pairs(tracks.df[c(15, 7:8)], lower.panel = NULL,
#       upper.panel = upper.panel)
# 
# pairs(tracks.df[c(15, 9:10)], lower.panel = NULL,
#       upper.panel = upper.panel)
# 
# pairs(tracks.df[c(15, 11:12)], lower.panel = NULL,
#       upper.panel = upper.panel)
```

Bargraph to show correlation popularity against continuous variables
```{r}
library(dplyr)
library(corrr)
library(ggplot2)

# Include on columns with continuous quantitative data
tracks_onlycontinuous.df <- subset(tracks.df, select = -c(lyrics, country, available_markets, key, id))

# Create bar chart of pearson correlation scores
tracks_onlycontinuous.df %>% correlate() %>% focus(popularity) %>%
  ggplot(aes(x = term, y = popularity)) +
  ggtitle("Correlation Between Popularity and Categories\n
    of Continuous Variables") +
    geom_bar(stat = "identity") +
    ylab("Pearson Correlation") +
    xlab("Variable") +
    ylim(-1, 1) +
    geom_col(aes(fill = popularity)) + 
    scale_fill_gradient2(low = "blue", 
                       high = "red", 
                       mid = "white",
                       midpoint = median(0))+
    theme(
      legend.position = "none",
      plot.title = element_text(size = 16, hjust = 0.5),
      axis.text.y = element_text(size = 12),
      axis.text.x = element_text(angle = 45, vjust = 1, 
        size = 12, hjust = 1),
      axis.title.y = element_text(size = 14),
      axis.title.x = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      panel.background = element_blank(),
      axis.ticks = element_blank()
    )
```
Conclusion: Nothing correlates strongly with popularity.

Popularity vs. Key (only categorical variable)
```{r}
# Key was left out of previous plot because it is a discrete variable

# Violin plot of keys vs. popularity
ggplot(tracks.df, aes(factor(key), popularity)) + 
  ggtitle("Correlation Between Popularity and Key", ) +
    ylab("Popularity") +
    xlab("Key") +
  geom_violin(fill = 'sienna1') + 
  geom_point(shape = 1, colour = "black", fill = "white", size = 0.01) +
  theme(legend.position = "none",
        panel.background = NULL,
        plot.title = element_text(size = 16, hjust = 0.5),
        axis.text.y = element_text(size = 12),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 14),
        axis.title.x = element_text(size = 14)
        )

#need key translations
```
Conclusion: No association of popularity with any keys.

boxplots comparing top and bottom 5% of songs in terms of popularity
```{r}
library(tidyr)
library(dplyr)

# Take top 5 most popular tracks in tracks.df
pop_95.df <- tracks.df %>%
  select(-c(lyrics, country, available_markets, key, tempo)) %>%
  filter(popularity > quantile(popularity, 0.95))

# Take bottom 5 least popular tracks in tracks.df
pop_5.df <- tracks.df %>%
  select(-c(lyrics, country, available_markets, key, tempo)) %>%
  filter(popularity < quantile(popularity, 0.05))

# Create column place that identifies each dataframe as either top 5 or bottom 5
pop_95.df$place <- 'top_5'
pop_5.df$place <- 'bottom_5'

# Combine the two data frames vertically
pop.df <- rbind(pop_95.df, pop_5.df)

# Graph boxplot
pop.df %>% 
  pivot_longer(., cols = c(acousticness, danceability, duration_scaled, energy, instrumentalness, liveness, loudness_scaled, speechiness, valence, popularity), names_to = "Var", values_to = "Val") %>%
  ggplot(aes(x = Var, y = Val, fill = place)) +
  ggtitle("Category Scores for Top 5% vs.\n
  Bottom 5% Most Popular Tracks") +
  geom_boxplot() +
  theme(
  plot.title = element_text(hjust = 0.5, size = 18),
  axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 10, hjust = 1),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.title = element_blank(),
  ) +
  scale_fill_manual(values=c('sienna1', 'mediumpurple1'))

```
Boxplot summary:  When comparing the bottom 5 vs. the top 5 most popular tracks, the two groups differ on most of the column categories.

Heatmap to show correlation between elements other than popularity

```{r}
library(reshape2)
library(ggplot2)

# subset tracks to include only continuous variables
tracks_onlycontinuous.df <- subset(tracks.df, select = -c(lyrics, country, available_markets, key, tempo, id))

# Create correlation table
tracks_cor <- round(cor(tracks_onlycontinuous.df, tracks_onlycontinuous.df), 2)

# Get upper triangle of the correlation matrix
get_upper_tri <- function(tracks_cor){
  tracks_cor[lower.tri(tracks_cor)]<- NA
  return(tracks_cor)
}
  
# Apply get_upper_tri to tracks_cor
upper_tri <- get_upper_tri(tracks_cor)

# Melt the correlation matrix
tracks_melt <- melt(upper_tri, na.rm = TRUE)

# Create a ggheatmap
ggplot(tracks_melt, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
    midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
ggtitle("Correlation Between Numerical Categories") +
geom_text(aes(Var2, Var1, label = value), color = "black", size = 3) +
theme(
  plot.title = element_text(hjust = 0.5),
  axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 10, hjust = 1),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
```

Correlations summary:

Energy and acousticness have a negative -0.72 correlation, loudness and energy have a 0.78 correlation, lyrics and instrumentalness have a 0.51, loudness and acousticness have a -0.58, and valence and danceability have a 0.52.

## STUDY OF THE MOST COMMON WORDS IN LYRICS
### MOST COMMON WORDS IN LYRICS
```{r}
#Install packages
#install.packages("viridis")

#Import Libraries
library(ggpubr)
library(Matrix)
library(qdapDictionaries)
library(readxl)

#Read stop words list, that contains prepositions, articles and pronouns to be removed from the lyrics
df.stop_words <- read_excel("D:\\Datos\\Documents\\Development\\R\\Data Munging\\Assignments\\Group Project II - Spotify Tracks\\stop_words.xlsx")
stop_words <- c(df.stop_words$words)

```

```{r}
#Create dataframe with only the data in lyrics, id and popularity columns
df.lyrics <- tracks.df %>% 
  select(lyrics, id, popularity)

#Preprocessing: lowercasing, smooshing contractions, removing punctuation, numbers, and getting rid of extra spaces
df.lyrics <- df.lyrics %>% 
  dplyr::mutate(
    clean_text = tolower(lyrics),
    clean_text = gsub("[']", "", clean_text), 
    clean_text = gsub("[[:punct:]]", " ", clean_text),
    clean_text = gsub('[[:digit:]]+', " ", clean_text),
    clean_text = gsub("[[:space:]]+", " ", clean_text),
    clean_text = trimws(clean_text))
```

```{r}
#Convert lyrics into a Document Term Matrix that counts the number of times that each words appear

text <- df.lyrics$clean_text
doc_id <- df.lyrics$id

#Convert the lyrics of each entry into a list of words (tokens)
tokns <- strsplit(text, " ", fixed=TRUE)

# Clean tokens before inserting in the matrix
for (i in 1:length(tokns)){
  current <- tokns[[i]]
  # Retrieve only the words contained within the English dictionary
  current <- current[current %in% GradyAugmented]
  # Retrieve only meaningful words (clean stop-words: prepositions, articles and pronouns)
  current <- current[!current %in% stop_words]
  # Replace into original list of tokens
  tokns[[i]] <- current 
}

#Create Document Term Matrix:
# Rows: Tracks
# Columns: Words
# Each position of the matrix [ij] contains how many times the word in column j appears in track of row i
vocab <- sort(unique(unlist(tokns)))
dtm <- matrix(data = 0L, 
              ncol = length(vocab), nrow = length(tokns),
              dimnames = list(doc_id, vocab) )
freqs <- lapply(tokns, table)
for (i in 1:length(freqs) ){
  doc <- freqs[[i]]
  dtm[i, names(doc)] <- as.integer(doc)
}

```
```{r}
#install.packages("RColorBrewer")
library(RColorBrewer)

#Count the total number of occurrences of each word
words_count <- colSums(dtm)
#Sort words per number of occurrences
words_count <- sort(words_count, decreasing=TRUE)

#Get the first 25 words (most occurrences)
top_25_words <- words_count[0:25]

# Plot the most frequent words contained in the songs
barplot(sort(top_25_words[1:20]), 
        ylab = "Words",
        xlab = "Words frequencies",
        xlim = c(0,1600),
        horiz = T,
        las = 1,
        names.arg = names(sort(top_25_words[1:20])),
        col = brewer.pal(20, "Pastel1"),
        main = "Top 25 words in songs",
        cex.names = 0.7)
```

```{r}
#Generate word-cloud graph

#install.packages("wordcloud")
#install.packages("RColorBrewer")
#install.packages("wordcloud2")
library(wordcloud2)
library(RColorBrewer)
library(wordcloud)

top_100_words <- words_count[0:100]
df.top_100_words <- data.frame(keyName=names(top_100_words), 
                               count=top_100_words, row.names=NULL)

wordcloud(words = df.top_100_words$keyName, freq = df.top_100_words$count, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
```


### MOST COMMON WORDS IN TOP 100 POPULAR SONGS
```{r}
#Arrange dataset according to popularity
df.popular <- df.lyrics %>% 
  arrange(desc(popularity))

#Get top 100 popular songs
df.popular_top_100.lyrics <- df.popular[1:100,] %>% 
  select(clean_text, id, popularity)
```

```{r}
#Convert lyrics into a matrix that counts the number of times that the words appear

text_pop <- df.popular_top_100.lyrics$clean_text
doc_id_pop <- df.popular_top_100.lyrics$id

#Convert the lyrics of each entry into a list of words (tokens)
tokns_pop <- strsplit(text_pop, " ", fixed=TRUE)

# Clean tokens
for (i in 1:length(tokns_pop)){
  current <- tokns_pop[[i]]
  # Retrieve only the words contained within the English dictionary
  current <- current[current %in% GradyAugmented]
  # Retrieve only meaningful words (clean stop-words: prepositions, articles and pronouns)
  current <- current[!current %in% stop_words]
  # Replace into original list of tokens
  tokns_pop[[i]] <- current 
}

vocab_pop <- sort(unique(unlist(tokns_pop)))
dtm_pop <- matrix(data = 0L, 
              ncol = length(vocab_pop), nrow = length(tokns_pop),
              dimnames = list(doc_id_pop, vocab_pop) )
freqs_pop <- lapply(tokns_pop, table)
for (i in 1:length(freqs_pop) ){
  doc <- freqs_pop[[i]]
  dtm_pop[i, names(doc)] <- as.integer(doc)
}
```

```{r}

words_count_pop <- colSums(dtm_pop)
words_count_pop <- sort(words_count_pop, decreasing=TRUE)
#head(words_count_pop)
top_100_words_pop <- words_count_pop[0:100]

# Plot the most frequent words contained in the most popular songs
barplot(sort(top_100_words_pop[1:20]), 
        ylab = "Words",
        xlab = "Words frequencies",
        xlim = c(0,500),
        horiz = T,
        las = 1,
        names.arg = names(sort(top_100_words_pop[1:20])),
        col = brewer.pal(20, "Pastel2"),
        main = "Top 20 words in most popular songs",
        cex.names = 0.7)
```

```{r}
#Generate word-cloud graph

df.top_100_words_pop <- data.frame(keyName=names(top_100_words_pop), 
                                   count=top_100_words_pop, row.names=NULL)

wordcloud(words = df.top_100_words_pop$keyName, freq = df.top_100_words_pop$count, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
```

